# Assignment

## General settings

```{r global_options, include=TRUE}
knitr::opts_chunk$set(fig.width=12, fig.height=8, fig.path='figure/',
                      echo=TRUE, warning=FALSE, message=FALSE)
```


## Loading and preprocessing the data

The data are loaded from the working directory and the variable "date" is transformed to class "Date"


``` {R,echo=TRUE}
Data <- read.csv("activity.csv")
Data$date <-as.Date(Data$date)
```

## What is the mean total number of steps taken per day?

1. Histogram of the total number of steps taken each day

```` {R}
sumsteps<-aggregate(steps~date,Data,sum,na.rm=T)
hist(sumsteps$steps, xlab="no. of steps",main="Steps per day",col="lightblue")
````

2. Calculating the mean and median of total steps taken per day

```{R}
mean(sumsteps$steps,na.rm = T)
median(sumsteps$steps,na.rm=T)
````

## What is the average daily activity pattern?

1. Number of steps taken averaged across all days and creating a series plot

```{R}
meansteps<- aggregate(steps~interval,Data,mean,na.rm=T)
plot(meansteps$interval,meansteps$steps,type = "l",xlab="Interval", ylab="No. of steps",main="Average of 53 observed days")
````

2. Interval which contains the maximum number of steps

```` {R}
max_step_int <- meansteps[meansteps$steps==max(meansteps$steps),1]
max_step_int
````

##Imputing missing values

1. Number of missing values

```` {R}
count_na <- sum(is.na(Data))
count_na
````

2. Strategy for filling in all of the missing values in the dataset and creating new dataset

We use the **Hmisc** package to impute the missing values and create an new dataset "Data2"

````{r,results="hide",warning=FALSE}
library(Hmisc)
set.seed(1234)
imp<-aregImpute(~steps+date+interval,Data)
imputation <- as.vector(imp$imputed$steps[,5])
Data2<-Data[order(Data$steps,na.last=F),]
Data2$steps[1:count_na] <-imputation
````

3. Histogram of the total number of steps taken each day and calculation of mean and median

````{r}
sumsteps_im <- aggregate(steps~date,Data2,sum)
hist(sumsteps_im$steps, xlab="no. of steps", main= "Steps per day",col="red")

mean(sumsteps_im$steps)
median(sumsteps_im$steps)
````

The average of steps taken slightly increases from **10766.19** (with missing values excluded) to **10780.49** (missing values imputed).The median of steps taken goes down from **10765.00** (missing values excluded) to **10600.00** (missing values imputed).

## Are there differences in activity patterns between weekdays and weekends?

1. Creating a new factor variable with two levels "weekday" and "weekend"

````{R}
Data2$wend <- (weekdays(Data2$date) == "Samstag") + (weekdays(Data2$date) == "Sonntag")
Data2$wend <- sub(1,"weekend",Data2$wend)
Data2$wend <- sub(0,"weekday",Data2$wend)
weekend<-subset(Data2,wend=="weekend")
weekday<-subset(Data2,wend=="weekday")
````


2. Time series plot of the average number of steps taken grouped by wekend and weekday.

````{R}
wend<-aggregate(steps~interval,weekend,mean) 
wday<-aggregate(steps~interval,weekday,mean)
plot(wend$interval, wend$steps,type="l",xlab="Interval", ylab="Average no. of steps", main="Weekend")
plot(wday$interval, wday$steps,type="l",xlab="Interval", ylab="Average no. of steps", main="Weekday")
````




